Nature or Nurture? Collisionless Evolution of Galactic Disc-Halo Systems

### 2019-07-19 Thesis/Paper_III.tex
Night before submission. Likely final pre-submission draft. Checked for spelling issues.

\chapter{Forming Vertical Structure in $\Lambda$CDM Discs}\label{ch:paper_iii}

>>>>

\newcommand{\kpch}{\ensuremath{h^{-1} \, \text{kpc}}}
\newcommand{\kpc}{\ensuremath{\text{kpc}}}
\newcommand{\solarmh}{\ensuremath{h^{-1} \, M_\odot}}
\newcommand{\solarm}{M_\odot}
\chapter{Forming Vertical Structure in $\Lambda$CDM Discs}\label{ch:paper_iii}
\textbf{This chapter contains a draft of a paper in preparation to be submitted to Monthly Notices of the Royal Astronomical Society. The working title for the paper is ``Vertical Structure and Kicked-Up Stars in $\Lambda$CDM Galaxies". }

\newpage
\section{Abstract}

We present and study vertical structure evolution twelve high resolution simulations of Milky Way stellar discs inserted into cosmological dark matter haloes. The simulations exhibit a wide variety of vertical structure formation arising from bar buckling, interactions with the host halo, and substructure. A systematic study of the effects of the disc's initial dynamical temperature, measured by Toomre $Q$, and the time of insertion on the disc properties is conducted. Inserting a disc into a more equilibrated halo appears to lead to quieter vertical structure evolution, and increasing the disc's internal dynamical temperature appears to have a non-trivial impact. Furthermore, we identify a population of stars that get kicked up to high latitudes in each simulation. Our results are consistent with the idea that a disc-satellite interaction origin of these structures requires a perturber in the region of $10^{11} \, \solarm$ in close proximity to the disc. A broader interaction with the triaxial host halo is much better at generating these higher latitude structures on short timescales.  As such, we determine that a wide variety of cosmological environments are able to explain low-latitude disc structures like Monoceros, A13, and TriAnd.

\vspace{1in}

\section{Introduction} \label{sec:introduction}
%Literature review about vertical structure.

% Observations tell us that the disc has vertical structure (Maybe say something about gas?)

Observations of the Milky Way and other spiral galaxies reveal an enormous range of vertical structure in stellar discs. Among the most prevalent is the presence of warps in the Milky Way's stellar and gas discs (see \citet{binney_1992} and \citet{Sellwood2013} for detailed reviews). Evidence for a warp in the Milky Way's \textsc{Hi} dates  back to the work of \citet{oort_1958}.  There is also evidence of a stellar warp, which roughly traces the \textsc{Hi} warp \citep{cox_1996, reyle_2009}. Beyond the Milky Way, extragalactic radio observations reveal how extremely common \textsc{Hi} warps are in observed edge-on spiral galaxies \citep[see][for example]{sancisi_1976,bosma_1991, garcia-ruiz_2002}.

Recent surveys have revealed other manifestations of vertical structure. Studies have focused on apparent vertical asymmetries in the thin disc in the Solar neighborhood in SDSS data \citep{widrow_2012_sdss, xu_2015}, LAMOST data \citep{carlin_2013_lamost, pearl_2017}, RAVE data \citep{williams_2013_rave, carrillo_2018_rave}, and Gaia data \citep{gaia_collab,bennet_2019_gaia}. These studies reveal both small scale corrugations as well as global bending and breathing waves in the stellar disc. Corrugations, mainly in the velocity structure, in extragalactic discs have also been observed \citep[for example]{matthews_2008,sanchez_2015}.

In addition to waves and corrugations, several stream-like structures appear to have disc-like dynamics. Conventionally, stellar streams are thought to originate from the tidal disruption of dwarf galaxies and globular clusters. Recently the idea that some streams could have disc origins has gained more traction. One example is the Monoceros ring \citep{newberg_2002, yanny_2003}, also called the Galactic Anticentre Stellar Structure (GASS) \citep{crane_2003, rocha-pinto_2003}. It is located at low latitudes towards the Galactic anti-centre at a Galactocentric distance of about $18\,\,\kpc$.  The coincidence in position and velocity of GASS and disc stars suggests that a disc origin may be a more likely explanation than a merger event. Similar structures have been found in the outer disc, most notably the Triangulum-Andromeda clouds (TriAnd) \citep{triand_discovery} and A13 \citep{a13_discovery}. This prompts a question about whether they too are associated with the disc \citep{johnston_kud_review}. In this hypothesis, the outer disc structures are simply rings in the disc with substantially different angular momentum. Although there is some speculation that these structures are associated with merger events, a disc origin is promising \citep{deason_2018,monoceros_disk_origin}.\\

Evidence for stream-like components in the thin discs of galaxies is not limited to the Milky Way. For instance, disc substructure has been observed in M31 \citep{richardson_2008_m31, dorman_2013_m31, bernard_2015_m31} and M33 \citep{mcconnachie_2006_m33, mcconnachie_2010_m33}. Due to the incredibly faint nature of low latitude structures like the ones seen in the Milky Way, direct observation of analogous structures in M31 is not yet feasible \citep{johnston_kud_review}.

%
In response to these and other observations, there is now a large concerted effort to explain vertical structure and low latitude features in galaxies like the Milky Way. This effort has primarily focused on the hypothesis that a single massive subhalo, such as the Sgr dSph \citep{ibata_discovery} is largely responsible for vertical structure in the Milky Way. Simulations of a single satellite encounter with an isolated galaxy illustrate the consequences of this hypothesis \citep[for example]{purcell2011,gomez_2013,widrow_2014,feldmann_2015, dlv_2015, donghia_2016, laporte_2016, laporte_2018, laporte_2018_b}. As far as vertical structure is concerned, the single satellite hypothesis appears to require a very massive, order $10^{11}\,\solarm$ Sgr dSph to explain low latitude streams in the outer disc.

Beyond a single satellite, other models attempt to create a fully realistic distribution of substructure to study its combined effect. For example, \citet{gauthier_2006} and \citet{chequers_2018} use a model which replaces a fraction of halo mass with substructure to study the evolution of the disc. \citet{chequers_2018} found that a plethora of vertical structure is generated when discs are embedded in this environment. However, these simulations also fail to create tilted rings of stripped stars.

These works suggest that a close encounter of with a massive satellite will excite vertical structure throughout the disc. While a strong local perturbation is one way to accomplish this, it is not necessarily the only way through a satellite encounter. Discs are not rigid bodies, and center of mass acceleration causes the disc to bend substantially on its own. Furthermore, these accelerations are ubiquitous in $\Lambda$CDM with every massive satellite infall. An experiment in \citet{sellwood_1996} gave a disc an axisymmetric, radially dependent displacement, and observed distinct axisymmetric flapping modes form in the disc. These perturbations were on the length scale of vertical structure expected from isolated subhalo encounters. As such, we might expect substantial vertical structure from any encounter yielding differential vertical acceleration.

% Global tides are another

The failure of satellite encounters to fully account for low latitude structure, namely A13 and TriAnd, in the Milky Way is not necessarily cause for concern. As \citet{binney_1992} points out in the context of warps, one does not necessarily need a close satellite encounter to excite vertical structure. The effect of global tides from disc-halo misalignment has been explored as a mechanism for exciting vertical structure and also for stripping stars from a thin disc. For example, \citet{hu_2016} embed discs in triaxial haloes. In their appendix, strong warps can be seen in discs which are substantially misaligned from their haloes. \citet{gomez_2017} found that vertical structure was ubiquitous in their hydrodynamical cosmological simulations, some of which did not have discs in a stable orientation to their host halo.

It is worth noting that disc-halo misalignment is not the only way to induce substantial global tides. A large LMC-mass object may be massive enough on its own to excite vertical structure, even on first infall. Indeed, as \citet{binney_1992} points out in the case of warps, a perpendicular component of torque on the disc is all that is required to excite the disc.

% Fully cosmological models

In this paper, we study how vertical structure evolves in a realistic $\Lambda$CDM environment, where potentially all of the above mechanisms are at play. Though this can be done using high resolution hydrodynamical simulations as in \citet{gomez_2017}, we use a disc insertion technique like the ones developed in \citet{berentzen_2006, debuhr_2012, ys_2015, bauer2018a, hu_2018}.

Our technique was successfully used to study bar formation in a cosmological context in \citet{bauer2018b}. Bar buckling is another mechanism for exciting vertical structure in stellar discs \citep{bar_buckling_echo}. When unstable bending waves are present at the centres of bar-forming discs, this gives rise to a time-dependent vertical potential. In turn, this excites new bending waves throughout the disc. This paper studies this important mechanism in more detail.

We demonstrate that a variety of halo environments could plausibly explain the Milky Way's vertical structure without the need for a very massive Sgr dSph. A number of the above effects are significant, and Monoceros-like structure appears ubiquitous. We also show that low latitude streams in the outer disc simply require a consistent perpendicular torque in the disc plane to be generated.

Our paper is then organized as follows. To motivate our simulation suite, we present two toy models in \S\ref{sec:overview} to serve as discussion points. In particular, they focus on the important effect of inertial acceleration of the disc.  The time evolution of key structural statistics in all models is studied in \S \ref{sec:evolution}. Focusing on the outer disc, we examine how stars can be stripped to high latitudes in \S \ref{sec:kud}. To connect the previous results to their halo environments, we examine several representative evolution scenarios in more detail in \S\ref{sec:disentangle}. These scenarios include a disc-halo misalignment dominated system, a substructure interaction dominated system, and a system which is somewhere in between. Finally, we discuss our key results in \S\ref{sec:discussion}.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/isolated_z_0_r_a.eps}
	\caption{Mean height, $\langle z \rangle$ of the disc from the toy model described in \S\ref{ssec:toy_model_1} as a function of cylindrical radius and time.} \label{fig:toy_model_1_mean_height}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/isolated_z_0_r_a_halo.eps}
	\caption{Mean height, $\langle z \rangle$ of the halo from the toy model described in \S\ref{ssec:toy_model_1} as a function of spherical radius and time.} \label{fig:toy_model_1_mean_height_halo}
\end{figure}

\section{Overview} \label{sec:overview}

In this section, we illustrate the factors driving vertical structure in stellar discs with two toy models.
\subsection{Quantifying Structure in Discs}
%\vspace{0.6in}

The analysis of simulated galaxies invariably involves computing averages of physical quantities, namely functions of the phase-space coordinates, $\textbf{r}$ and $\textbf{v}$. For some observable function,  $\Theta(\textbf{r},\textbf{v})$, the average in a region $S$ is,
\begin{eqnarray}
\langle\Theta\rangle_S &=&\frac{1}{M_S} \int_S \text{d}^3 \textbf{r} \text{d}^3\textbf{v} f(\textbf{r}, \textbf{v}) \Theta(\textbf{r}, \textbf{v})\\
&\approx& \frac{1}{M_S} \sum_{j \in S} m_j \Theta(\textbf{r}_j, \textbf{v}_j) \label{eq:z_statistic}
\end{eqnarray}
where $M_S$ is the mass in $S$, $f(\textbf{r},\textbf{v})$ is the phase space distribution function, and $m_j$ is the mass of the $j$-th particle in $S$. For stellar discs, we are often interested in the Fourier series representation of $\Theta$ to quantify $m$-fold angular symmetries in some annulus with mean radius, $R_S,$
\begin{equation}
\langle\Theta(R_S,m)\rangle_S = \frac{1}{M_S} \sum_{j \in S} e^{i m \phi_j} m_j \Theta(\textbf{r}, \textbf{v})
\end{equation}
where $m$ is the degree of symmetry, and $\phi_j$ is the azimuthal angle of particle $j$. When $\Theta = z^n$, we call the average $Z_m^n$. We make extensive use of this form for this paper.

\subsection{Toy Model: Acceleration Along the Symmetry Axis} \label{ssec:toy_model_1}

For our first simulation, we study the vertical structure that is excited when a dark matter subhalo falls along the disc symmetry axis from a height of 100 kpc.  We generate the initial conditions using  the \textsc{GalactICS} code \citep{KGGalactICSReference,WPDGalactICSReference}, and simulate the system using \textsc{Gadget-3}, a privately released version of the popular \textsc{Gadget-2} code \citep{GadgetCodePaper}.

% This experiment is similar to Sellwood 1996...
The disc has an exponential surface density with scale length $R_d=3.7\,\kpc$ and mass $M_d=4\times10^{10} \,\solarm$. The halo is taken to be NFW, with scale length $R_h=33.2\,\kpc$, virial mass $M_h=1.1\times10^{12}$, and concentration $c=6.7$\footnote{The parameters for the halo are taken to be a sphericalized version of simulation B.Late in Table \ref{tab:sims}. This simulation will be described in more detail in \S\ref{sec:description}.}. We use $10^6$ halo particles and $3.5 \times 10^{6}$ disc particles. The subhalo is also assumed to be NFW with scale length $R_h=1.70\,\kpc$, virial mass $M_h=2.7\times10^{10}$, and concentration $c=38.1$\footnote{ The parameters for the substructure are given by B.Late.a in Table \ref{tab:substructure}.}. The subhalo is given a height of $z=100\,\kpc$ and zero velocity.

A disc-centered mean height map for disc particles is shown in Fig. \ref{fig:toy_model_1_mean_height}. Initially the disc remains relatively unperturbed; the subhalo is far enough away to exert a somewhat uniform acceleration across the disc. Only in the very outer disc do we see very slight disc flapping. As the subhalo makes its first pericentre, the inner disc is dragged toward the subhalo more strongly, causing a strong lag in the outer disc. After the subhalo has passed through the disc, the pattern reverses as the center of the disc is pulled back down. The resultant lag of the outer disc then shows up as a positive signal.

The oscillation pattern seen in Fig. \ref{fig:toy_model_1_mean_height} is not stationary in time, but rather becomes more complicated as time goes on. Immediately after the displacement first changes signs, the pattern starts to propagate outwards more slowly as evidenced by decreasing (increasing magnitude) slope of the ridges. The mean height shows oscillating single fixed-point standing wave behaviour, at some points (around 1 Gyr) appearing to gain another node.  As the system equilibrates, the $m=0$ disturbances are pushed further to the outer disc, but the magnitude of the perturbations remains high.

Qualitatively similar behaviour can be seen in the halo. Fig. \ref{fig:toy_model_1_mean_height_halo} shows the mean height of spherical shells in the halo. It paints a similar picture to the disc; the inner shells are dragged with the disc as the substructure pulls on the inner disc-halo system more strongly, while the outer halo lags the oscillations. The shells in the outer halo are displaced with longer periods of oscillation than in annuli in the outer disc, indicating a more complicated interaction with the substructure and inner disc-halo system. The halo develops multiple nodes in mean height by the end of the simulation, suggesting a similar standing wave-like phenomenon is occurring in the halo.

What is interesting about this experiment is the very long-lived nature of the flapping in the disc. Vertical instabilities are generally expected to phase mix, causing the disc to thicken. The flapping in our disc persists for longer than a dynamical time. In fact, a similar observation was made in an experiment performed by \citet{sellwood_1996} where a disc with similar properties was given an initial bulk displacement that varied as a function of radius. While other bending behaviour is rapidly suppressed, a flapping motion remained in the perturbed disc. Qualitatively, the results of \citet{sellwood_1996} are similar in magnitude to our own, suggesting that flapping naturally arises from near-axisymmetric perturbations of the disc.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/isolated_orbits_two_fits.eps}
	\caption{The orbit of the disc (blue), halo (red), and subhalo(green) peak densities of the toy model described in \S\ref{ssec:toy_model_2}. The dashed black lines show the fits described in the text.} \label{fig:toy_model_2_orbits}
\end{figure}

\subsection{Toy Model: Massive Subhalo Interaction} \label{ssec:toy_model_2}
\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/b_late_light_subhalo_z_0_r_a.eps}
	\caption{Mean height, $\langle z \rangle$ of the disc from the toy model described in \S\ref{ssec:toy_model_2} as a function of radius and time.} \label{fig:toy_model_2_mean_height}
\end{figure}

Here, we present a more complicated satellite encounter. The orbit is inspired by a satellite in one of our cosmological simulations, and shows a massive satellite on a close orbit with the disc, passing through the midplane several times. The parameters for the disc, halo, and subhalo are the same as in \S\ref{ssec:toy_model_1}.

We plot the $z$ coordinate of the subhalo, disc, and halo peak densities in Fig. \ref{fig:toy_model_2_orbits}.
Initially, the subhalo is far from the disc at over 100 kpc, and about 20 kpc above the disc in $z$. We see the disc and halo peak densities track each other almost exactly over the course of the simulation, consistent with Fig. \ref{fig:toy_model_1_mean_height} and Fig. \ref{fig:toy_model_1_mean_height_halo}. In some short frame of time, we suggest that this system can be viewed as a two body problem between the disc-halo system and subhalo. That is, we have motion described by,
\begin{equation}
\frac{M_{dh}\textbf{r}_{dh} +  + M_{sh} \textbf{r}_{sh}}{M_{dh} + M_{sh}} = \textbf{r}_{c,0} + \textbf{v}_{c,0} t
\end{equation}
where $M_{dh}$ is the implied two-body disc-halo mass, $M_{sh}$ is the implied two-body subhalo mass, $\textbf{r}_{dh} = (M_d \textbf{r}_d + M_h \textbf{r}_h)/M_{dh}$, $\textbf{r}_d$ is the disc peak density position, $\textbf{r}_h$ is the halo peak density position, $\textbf{r}_{sh}$ is the subhalo peak density position, $\textbf{r}_{c,0}$ is the initial implied center of mass, and $\textbf{v}_{c,0}$ is the implied center of mass velocity.

We can find the density modes from the particle distributions each time very easily. However, $M_{dh}$, $M_{sh}$, $\textbf{r}_{c,0}$, and $\textbf{v}_{c,0}$ are unknowns. We maximize the likelihood,
\begin{equation}
\ln \mathcal{L} = \left\vert \frac{M_{dh}\textbf{r}_{dh} +  + M_{sh} \textbf{r}_{sh}}{M_{dh} + M_{sh}} - \textbf{r}_{c,0} + \textbf{v}_{c,0} t \right\vert^2 \, \text{kpc}^{-2}
\end{equation}
in two 1.0 Gyr windows  centered on 0.5 Gyr and 2.0 Gyr, respectively. These windows are chosen since they encompass the time before pericentric passages at 0.85 Gyr and 3.0 Gyr, when the subhalo is losing the most mass. The idea is that during pericentric passages, we expect the system to be poorly described by a two-body fit. The two center of mass motion lines, $\textbf{r}_{c,0} + \textbf{v}_{c,0} t$ are shown as the solid black and dashed black lines in Fig. \ref{fig:toy_model_2_orbits}, respectively. We imply effective two-body mass ratios of 16.3:1 and 85.0:1, respectively, for disc-halo:subhalo masses.

The likelihood represents the mean distance deviation from our fit. For the two windows centered at 0.5 Gyr and 2.0 Gyr, we have the average root-log-likelihood $(N^{-1}\ln\mathcal{L})^{1/2}= 0.19$ in the first window and  $(N^{-1}\ln\mathcal{L})^{0.5} = 0.29$ in the second, where $N$ is the number of observations in the window. We thus infer small deviations from a two body problem when the subhalos are not at pericenter, implying that this model describes the data well.  We submit this technique as a way to quantify the effect of inertial acceleration in isolated satellite encounters. Although the drift velocity of the system changes as the satellite sheds mass, the picture is clear: the disc is placed on an oscillating orbit in $z$, and is no doubt subject to differential vertical acceleration. Despite this acceleration, the disc axis as measured by its moment of inertia tensor is relatively undeflected, being constrained to less than $2^{\circ}$ in the first 4 Gyr.

We see the orbit's effect on the mean height of the disc in Fig. \ref{fig:toy_model_2_mean_height}. At the beginning of the simulation, although the amplitude of the oscillations is lower, we see qualitatively similar behaviour to the first toy model. There is no obvious mean height bump when the substructure is at pericenter, indicating that passages through the disc plane are less substantial for generating $m=0$ corrugations than the overall two-body orbit. We expect that any subhalo capable of creating two-body motion of the disc along the disc axis will generate substantial, several hundred parsec corrugations.

\section{Description of Experiments} \label{sec:description}

In this section, we proceed to generalize to a fully cosmological environment.

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/rotation_curve.eps}
	\caption{The rotation curves for all considered models. The same disc is used in all runs, and its rotation curve contribution shown by the black dashed line. The halo contributions  for the Early and Warm subsuites are shown in the top panel as dot-dashed lines, with the total rotation curves being solid. The bottom panel shows the same for the Late subsuite.} \label{fig:rotation_curves}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/timescale.eps}\caption{The timescale of torque acting on a ring for all models considered in this paper. Colours are the same as in Fig. 5. Line types are Early (solid), Warm (dot-dashed), and Late (dashed).} \label{fig:ratio_freqs}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{../figures/substructure_r_vs_t.eps}
	\caption{The orbits of the two most massive structures at initialization for the fiducial suite (top panel) and the late insertion suite (bottom). These are the substructures listed as the top two entries for each simulation in Table \ref{tab:substructure}. The red line denotes $20\, \kpch = 29.5 \, \kpc$. } \label{fig:substructure_orbits}
\end{figure}

\subsection{Disc Insertion Technique} \label{ssec:disc_insertio}

Our disc insertion scheme is described in \citet{bauer2018a}, and
it expands upon method introduced by
\citet{BerentzenShlosmanStellarDisks}, \citet{debuhr_2012}, and
\citet{ys_2015}. Our scheme uses an iterative procedure to
initialize an axisymmetric, smooth disc in a cosmological halo, generally clumpy and triaxial. 
The first step is to run a pure dark matter
simulation and identify a suitable halo.  We then rerun the simulation with a rigid disc potential which 
grows from an initial scale factor, $a_g$, to a final scale factor, $a_l$.
When the simulation reaches $a_l$,
the rigid disc is replaced by an N-body system and the ``live'' disc-halo
system is evolved to the present epoch.

For our pure dark matter simulation, we implement the zoom-in
technique of \citet{KatzQuasarZoom} and \citet{NavarroWhiteZoom}.
Selecting for Milky Way-like halos with no major mergers, we use the
results of \cite{onorbe_etal_2014} to ensure no low-resolution particle contamination.
We choose cosmological parameters based on the results
from Planck 2013 \citep{planck_2014} with $H_0=67.9\,{\rm
  km\,s}^{-1}\,{\rm kpc}^{-1}$, $\Omega_b = 0.0481$, $\Omega_0 =
0.306$, $\Omega_\Lambda = 0.694$, $\sigma_8 = 0.827$, and $n_s =
0.962$.  N-body initial conditions for the dark matter particles are
generated with the \textsc{music} code \citep{music}.

%We select a suitably-sized halo for a Milky Way-like galaxy, namely
%one with a $z=0$ mass of $1.23\times 10^6\,h^{-1} M_\odot$
%that comprises $10^6$.

During its growth phase from $a_g=0.25$ to $a_l=0.5$\footnote{For reference, $a=0.5$ is 5.9 Gyr after the Big Bang, and $a=0.75$ is 10.0 Gyr after the Big Bang. The age of this Universe at present day is 13.8 Gyr.}, the disc is treated
as a rigid body whose orientation and center-of-mass position evolve
according to the standard equations of rigid body dynamics. \citet{bauer2018a} found that
the rigid disc assumption is actually quite true to the orientation
and center of mass evolution of a pure stellar disc. At
$a_l$, we swap a live disc for the rigid one using a modified version of
\textsc{GalactICS} \citep{KGGalactICSReference,WPDGalactICSReference}, which generates a
three-integral DF disc in the best axisymmetric approximation to the
halo \citet{bauer2018a}.

\subsection{Numerical Experiments} \label{ssec:numerical_experiments}

%We apply the technique described in \S \ref{ssec:disc_insertio} to understand a possible disc origin of low-latitude streams, and more generally how vertical structure evolves in pure stellar discs. In particular, we are interested in understanding how, in a cosmological simulation, stars are kicked up out of the disc. They comprise a component, if not all member stars of low-latitude streams. We discuss useful definitions of this population subsequently in \S\ref{ssec:def}. Such kicked-up disc stars (KUD stars) may be useful probes of the Galactic potential if we can understand how their orbits were altered.

We present a concise description of our cosmological models in Table \ref{tab:sims}. The haloes are chosen such that the total circular speed curves are similar to the Milky Way's. One of the haloes is considered by \cite{bauer2018b}, and the other three, A, B, and C, are previously unpublished.

Fig. \ref{fig:rotation_curves} shows the circular speed curves for all disc-halo combinations. The top panel shows these for our fiducial suite of simulations, which has $a_l=0.5$. This suite of simulations is denoted for halo X as X.Early. We also run a suite of simulations with $a_l=0.75$, and the corresponding circular speed curves are shown in the bottom panel. This suite is denoted as X.Late. We see a submaximal disc in all cases, and a relatively small spread in circular velocities in the inner disc. In particular, the ratio of the square of the disc contribution to the rotation curve and square of the total rotation curve measured at $2.2\, R_d$, $\left. V_d^2/V_{tot}^2 \right \vert_{R=2.2\,R_d}$, ranges between 0.37 and 0.46. These values are far below the lower bound for a maximal disc as advocated by \citet{sackett_1997}.

One of the benefits of disc insertion techniques is the ability to modify disc properties in the same halo. We exploit this in our third suite denoted as X.Warm. In this suite, we have increased the disc's dynamical temperature measured by Toomre $Q$ \citep{toomre_q}.

Although we explore all of our models in \S\ref{sec:evolution} and \S\ref{sec:kud}, we identify three key simulations for further study in \S\ref{sec:disentangle}.
\begin{enumerate}
\item \textbf{Strong Subhalo Interaction:} B.Late has a close interaction with two subhaloes, but no substantial disc-halo misalignment. One of these subhaloes passes twice through the plane of the disc, one time interior to 30 kpc.
\item \textbf{Disc-Halo Misalignment:} C.Early presents a case with no massive subhalo interactions, but a substantial misalignment from its host halo.
\item \textbf{Distant Massive Subhaloes:} A.Early is relatively aligned to its host halo, but its most massive substructures do not interact closely with the disc.
\end{enumerate}
These simulations represent three archetypal halo environments, which we explore in more detail in subsequent sections.

\begin{sidewaystable}
\centering
\caption{Table of simulations run in the main experiment. $M_h$ is the virial mass of the host halo, $R_h$ is the NFW scale length of the host halo, $c$ is the NFW concentration parameter of the host halo, $a_l$ is the live disc insertion scale factor, $\delta \Theta_{h,20}$ is the halo misalignment angle for the 20 kpc shell, $N_d$ is the number of disc particles,  $Q$ is the Toomre stability parameter of the disc at initialization, and $f_{kud}$ is the fraction of disc mass that gets classified as KUD stars measured at present day with $\eta=8$.} \label{tab:sims}
\begin{tabular}{l l l l l l l l l l}
\hline
       & $M_h\,\,(\solarm)$ & $R_h\,\,(\kpc)$ & $c$ & $a_l$ &  $N_d$ & $Q$ & $\cos\delta \Theta_{h,20}$ & $f_{kud}$\\
\hline
bauer2018.Early   & $1.0 \times 10^{12}$ & 15.9 &  9.84 &   0.50 & $1.0 \times 10^6$ & 1.1 & 0.16 & $1.2 \times 10^{-2}$\\
bauer2018.Warm    & $1.0 \times 10^{12}$ & 15.9 &  9.84 &   0.50 & $1.0 \times 10^6$ & 1.1 & 0.16 & $1.1 \times 10^{-2}$\\
bauer2018.Late    & $1.2 \times 10^{12}$ & 15.2 &  15.4 &   0.75 & $1.0 \times 10^6$ & 1.8 & 0.93 &$4.2 \times 10^{-3}$\\
A.Early           & $9.4 \times 10^{11}$ & 17.8 &  8.56 &   0.50 & $3.5 \times 10^6$ & 1.3 & 0.97 &$3.7 \times 10^{-3}$\\
A.Warm            & $9.4 \times 10^{11}$ & 17.8 &  8.56 &   0.50 & $3.5 \times 10^6$ & 1.8 & 0.97 & $1.6 \times 10^{-3}$\\
A.Late            & $1.1 \times 10^{12}$ & 18.5 &  12.1 &   0.75 & $3.5 \times 10^6$ & 1.3 & 0.98 & $6.8 \times 10^{-5}$\\
B.Early           & $5.9 \times 10^{11}$ & 10.4 &  12.6 &   0.50 & $3.5 \times 10^6$ & 1.3 & 0.91 & $8.5 \times 10^{-4}$\\
B.Warm            & $5.9 \times 10^{11}$ & 10.4 &  12.6 &   0.50 & $3.5 \times 10^6$ & 1.8 & 0.91 & $1.2 \times 10^{-3}$\\
B.Late            & $1.1 \times 10^{12}$ & 33.2 &  6.70 &   0.75 & $3.5 \times 10^6$ & 1.3 & 0.99 & $9.2 \times 10^{-4}$\\
C.Early           & $4.3 \times 10^{11}$ & 8.33 &  14.1 &   0.50 & $3.5 \times 10^6$ & 1.3 & 0.84 & $5.3 \times 10^{-3}$\\
C.Warm            & $4.3 \times 10^{11}$ & 8.33 &  14.1 &   0.50 & $3.5 \times 10^6$ & 1.8 & 0.84 & $5.6 \times 10^{-3}$\\
C.Late            & $6.7 \times 10^{11}$ & 13.6 &  14.0 &   0.75 & $3.5 \times 10^6$ & 1.3 & 0.92 & $1.0 \times 10^{-4}$\\
\hline
\end{tabular}

\end{sidewaystable}

\subsection{Halo Environments} \label{ssec:halo_env}

%As mentioned in \S\ref{ssec:numerical_experiments}, the haloes we consider are meant to resemble the Milky Way. In Table \ref{tab:sims}, all of the simulations we consider are listed.

%In addition to basic properties like halo mass, $M_h$, and scale radius, $R_h$, we also want to know how equilibrated the halo is with the disc. Discs will tumble in response to an external specific torque, $\boldsymbol \tau$, induced by the non-equilibrated halo. This invariably produces vertical structure.

Here, we further quantify our halo environments in terms of the effects we expect their smooth distributions and subhalos to have on the disc. For the former of these, we examine the following timescale,
\begin{equation}
T_\tau = \frac{L(R)}{\tau_\perp(R)},
\end{equation}
where $L(R)$ is the magnitude of the angular momentum in a ring centered at radius $R$, and $\tau_\perp(R)$ is the cylindrically averaged perpendicular torque in an annulus centered at $R$. This timescale is plotted as a function of radius at initialization for our models in Fig. \ref{fig:ratio_freqs}. This lower the timescale, the more the initial halo configuration contributes to the relevant torque. For example, we would interpret the simulation C.Early as being substantially non-equilibrated at initialization. Generally speaking, the more misaligned disc-halo systems have shorter torque action timescales. The effect of having a warmer disc is most pronounced in the inner regions, where the increased angular momentum from asymmetric drift appears to increase $T_\tau$.

%In a non-equilibrium state, shells of the halo may be misaligned from others at different radii. This creates a torquing effect on the disc and causes it to reorient (see \cite{DubinskiKuijkenRigidDisks}, for example). Torques present an opportunity for the angular momenta of orbits to change, and may result in the generation of KUD stars. Fig. \ref{fig:disk_torques} shows the specific torque as a function of radius for all models. In the most misaligned simulations, bauer2018b.Early/Warm, bauer2018b.Late, and C.Early/Warm, there is an obvious torque signature in the intermediate parts of the disc.

%he relative tilt of the halo principal axes compared to the halo interior for all runs. The model considered in \citet{bauer2018b}, bauer2018b.Early stands out the most. It shows substantial misalignment at all radii. C.Early also shows substantial misalignment, but the other simulations seem to have mostly aligned short axes.\

Subhalo interactions pose another potential source of perpendicular torque. We expect the timescales over which these events act to be shorter, but we can understand them by looking at the subhalo orbits directly. Fig. \ref{fig:substructure_orbits} shows the two most massive substructures in a plot of $r$ vs $t$ for all runs. A list of the top five most massive (at initialization) substructures' masses and scale radii is presented in Table \ref{tab:substructure}. The orbits present a wide variety of behaviour. For example, the subhaloes in bauer2018b.Early live most of their lives in the outer halo on roughly tangential orbits. Opposite to this are radial orbits like the one of C.Late.b, for example. Of course, most orbits exist somewhere in between on this scale.
